---
title: "Lipidomics Baseline Analysis"
subtitle: "Data preprocessing"
author: "Sofie Olund Villumsen"
date: "31/8/2021"
output: 
  html_document:
    toc: true
    toc_depth: 2
---
R markdown doc analyzing and exploring lipidomics data in combination with clinical from the type 2 diabetes overfeeding study of 48 men with mean age of 38 years.

# Initial settings

- Inspect data format 
- Check for columns with zero or near zero variance
- Check for missing values
- Check for outliers
- Impute missing values and outliers with K-nearest neighbor
- Check normality (density plots and Kolmogorov-Smirnov test)
- Transform and scale data (log2 and zscore/autoscale)


```{r echo=FALSE}
# Clear workspace
rm(list = ls()) 
```

Load packages
```{r message=FALSE}
library(tidyverse)
library(caret) # Function nearzerovar
library(impute) # Imputing data with KNN
library(DataExplorer) # http://boxuancui.github.io/DataExplorer/
```

Load data
```{r message=FALSE}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie/Baseline_study/Lipidomics"
setwd(path)

lipid_data <- read_csv(paste(path, "data/01_lipidomics_data.csv", sep="/")) 
clinical_data <- read_csv(paste(path, "data/01_clinical_data.csv", sep="/"))
```

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------


# Tidy and combine lipidomics and clinical data

```{r echo=FALSE}
# Lipidomics data cleanup
lipid_data_tidy <- lipid_data %>% 
  separate(Sample_ID, into = c("Sample_ID", "Label"), sep = "(?<=[0-9])(?=[a-z])") %>% 
  mutate(Label = str_to_upper(Label), 
         Sample_ID = case_when(nchar(Sample_ID) == 2 ~ as.character(paste0("0", Sample_ID)))) %>% 
  select(everything(), -...282)
```


```{r}
# Change lipid names to accomodate the correct nomenclature XX(XX:XX)
lipid_names <- c()

for (lipid in colnames(lipid_data_tidy)) {
  if (str_detect(lipid,"Cer")) {
    cer <- gsub('^(\\S+)\\([d](\\d[0-9]*):(\\d[0-9]*)', '\\1-d(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, cer)
    
  } else if (str_detect(lipid,"SM")) {
    sm <- gsub('^(\\S+)\\([d](\\d[0-9]*):(\\d[0-9]*)\\)', '\\1-d(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, sm)
    
  } else if (str_detect(lipid,"TG|PG|PA|FA|MePE|CE|DG")) {
    lipid_names <- c(lipid_names, lipid)
    
  } else if (str_detect(lipid,"LPC\\(0:0/O-18:1\\)")) {
    lpc_special <- "LPC-O(18:1)"    
    lipid_names <- c(lipid_names, lpc_special)
    
  } else if (str_detect(lipid,"^LPC\\(0")) {
    lpc_1 <- gsub('LPC\\S+/(\\d[0-9]*):(\\d[0-9]*)\\)', 'LPC(\\1:\\2)', lipid)    
    lipid_names <- c(lipid_names, lpc_1)
    
  } else if (str_detect(lipid,"^LPC\\(\\d+")) {
    lpc_2 <- gsub('LPC\\((\\d[0-9]*):(\\d[0-9]*)', 'LPC(\\1:\\2)', lipid)    
    lipid_names <- c(lipid_names, lpc_2)
    
  } else if (str_detect(lipid,"^LPC\\(O")) {
    lpc_o <- gsub('^(\\S+)\\(O-(\\d[0-9]*):(\\d[0-9]*)', '\\1-O(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, lpc_o)
    
  } else if (str_detect(lipid,"^LPC\\(P")) {
    lpc_p <- gsub('^(\\S+)\\(P-(\\d[0-9]*):(\\d[0-9]*)\\)', '\\1-P(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, lpc_p)
    
  } else if (str_detect(lipid,"^PC\\(O")) {
    pc_o <- gsub('^(\\S+)\\(O-(\\d[0-9]*):(\\d[0-9]*)\\)', '\\1-O(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, pc_o)
    
  } else if (str_detect(lipid,"^PC\\(P")) {
    pc_p <- gsub('^(\\S+)\\(P-(\\d[0-9]*):(\\d[0-9]*)\\)', '\\1-P(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, pc_p)
    
  } else if (str_detect(lipid,"^PC\\(")) {
    pc <- gsub('^(\\S+)\\((\\d[0-9]{2})(\\d[0-9]*)\\)', '\\1-d(\\2\\3', lipid)
    lipid_names <- c(lipid_names, pc)
    
  } else if (str_detect(lipid,"^LPE\\(P")) {
    lpe_p <- gsub('^(\\S+)\\(P-(\\d[0-9]*):(\\d[0-9]*)\\)', '\\1-P(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, lpe_p)
    
  } else if (str_detect(lipid,"LPE\\(0:0/20:1\\)")) {
    lpe_special <- "LPE(21:1)"    
    lipid_names <- c(lipid_names, lpe_special)
    
  } else if (str_detect(lipid,"^LPE\\(")) {
    lpe <- gsub('^(\\S+)\\((\\d[0-9]*):(\\d[0-9]*)[/|\\)]\\S?', '\\1(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, lpe)
    
  } else if (str_detect(lipid,"^PE\\(O")) {
    pe_o <- gsub('^(\\S+)\\(O-(\\d[0-9]*):(\\d[0-9]*)\\)', '\\1-O(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, pe_o)
    
  } else if (str_detect(lipid,"^PE\\(P")) {
    pe_p <- gsub('^(\\S+)\\(P-(\\d[0-9]*):(\\d[0-9]*)\\)', '\\1-P(\\2:\\3)', lipid)
    lipid_names <- c(lipid_names, pe_p)
    
  } else if (str_detect(lipid,"^PE\\(")) {
    pe <- gsub('^(\\S+)\\((\\d[0-9]{2})(\\d[0-9]*)\\)', '\\1-d(\\2\\3', lipid)
    lipid_names <- c(lipid_names, pe)

  } else if (str_detect(lipid,"Sample_ID|Label")) {
    next
  } 
}

lipid_data_clean <- lipid_data_tidy
colnames(lipid_data_clean) <- c("Sample_ID","Label",lipid_names)
```




```{r echo=FALSE}
# Clinical data cleanup
clinical_data_clean <- clinical_data %>% 
  unite("bwcat_NAFLD", c(bwcat,NAFLD), remove = FALSE) %>% 
  filter(visit != "M") 
```

Combine cleaned data frames
```{r echo=FALSE}
combined_data <- left_join(lipid_data_clean, clinical_data_clean,  by = c("Label" = "visit", "Sample_ID" = "id")) %>% 
  select(Sample_ID, bwcat, randomization, Label, NAFLD, everything()) #%>% 
combined_data
```


---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------



# Explore data

## Variance check (remove cols with zero or near zero values)
```{r echo=FALSE}
nearZeroVar(lipid_data_clean[, 3:ncol(lipid_data_clean)], saveMetrics = TRUE) %>% 
  filter(nzv == TRUE)
```
**Comments:** No variance data, thus no need for removing any variables.


## Missing values in combined data
```{r echo=FALSE}
missing_lip <- profile_missing(lipid_data_clean) %>% arrange(desc(pct_missing))
missing_lip
```
**Comments:** 11 missing value in the data set (Cer(d43:1), PA(44:1), PA(44:2), PG(34:1), TG(58:12), TG(42:2), TG(40:1))


## Identify negative values in metaboloics data
```{r echo=FALSE}
minvals <- as.data.frame(apply(lipid_data_clean,2,min))
```
**Comments:** From the list of min vals, we see that the minimum values for each metabolite differs a lot and no lipids have negative concentration, however some have missing values ((PG(34:1),Cer(d43:1),TG(58:12),TG(42:2),TG(40:1),PA(44:1) and PA(44:2)). Missing values  will be imputed. 


## Outlier overview via boxplot
```{r echo=FALSE}
# Pivot data frame 
pivot_lipid_data <- lipid_data_clean %>% 
  pivot_longer(cols = -c(Sample_ID, Label), 
               names_to = "Lipid", 
               values_to = "Concentrations")

# Boxplot of raw metabolite concentrations
lipid_plot <- pivot_lipid_data %>% 
  ggplot(mapping = aes(x = Concentrations, y = Lipid)) + 
  geom_boxplot(alpha = 0.5) + 
  geom_point(position = position_jitter(seed = 0.4, width = 0.02), size = 0.8) +
  coord_flip() + 
  labs(x = 'Raw lipid concentrations', title = "Distribution of lipid concentrations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.5))
lipid_plot
```
**Comments:** from this plot it is clear that the lipids follow different ranges and that outliers do occur.


## Outliers to NAN
```{r}
lip_names <- lipid_data_clean %>% 
  select(everything(), -c(Sample_ID, Label))

lip_df <- lipid_data_clean %>% 
  unite("Label_SampleID", c(Label,Sample_ID), remove = TRUE) 

df_rm_outlier <- data.frame(lip_df$Label_SampleID)

# Loop for identifying outliers and replacing them with NA values
for (lip in colnames(lip_names)) {
  # Identify outliers
  x_new <- filter(lip_df, lip_df[[lip]] %in% boxplot(lip_df[[lip]], plot = FALSE)$out)
  
  # Replace the outlier values with NA
  lipid <- replace(lip_df[[lip]], lip_df[[lip]] %in% x_new[[lip]], values = NA)
  
  # Combine in data frame
  df_rm_outlier <- cbind(df_rm_outlier, data.frame(lipid))
}

colnames(df_rm_outlier) <- colnames(lip_df)
```


```{r}
table(is.na(df_rm_outlier))
sapply(df_rm_outlier, function(x) sum(is.na(x)))
```



---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------


## Impute data with k-nearest neighbor
```{r}
# Convert data frame from numeric to factor
indx <- sapply(df_rm_outlier, is.numeric)
df_rm_outlier[indx] <- lapply(df_rm_outlier[indx], function(df_rm_outlier) as.factor(as.character(df_rm_outlier)))
df_rm_outlier <- df_rm_outlier[2:280]

# Set random seed 
if(exists(".Random.seed")) rm(.Random.seed)

# Impute with KNN function
df_rm_outlier_imputed <- impute.knn(as.matrix(df_rm_outlier))

lip_df_imputed <- as_tibble(df_rm_outlier_imputed$data) %>% 
  mutate(Sample_ID = lipid_data_clean$Sample_ID,
         Label = lipid_data_clean$Label) %>% 
  select(Sample_ID, Label, everything())

```




## Impute
```{r echo=FALSE}
# List of lipids containing missing values
#miss_lip <- as.character(subset(missing_lip$feature, missing_lip$num_missing > 0))

#df_mean_total <- data.frame()

# Loop calculating mean value of lipids with missing values, creating a new df with mean values
#for (lipid in miss_lip) {
 # cal_mean <- lipid_data_clean[[lipid]] %>% 
  #  mean(na.rm = TRUE)
  #df_mean <- data.frame(lipid, cal_mean)
  #df_mean_total <- rbind(df_mean_total, df_mean)
#}

#df_mean_total

# Impute with values obtained
#lipid_data_clean_impt <- lipid_data_clean %>% 
 # mutate(`Cer-d(43:1))` = case_when(is.na(`Cer-d(43:1))`) ~ df_mean_total$cal_mean[1], TRUE ~ `Cer-d(43:1))`),
  #       `PA(44:1)` = case_when(is.na(`PA(44:1)`) ~ df_mean_total$cal_mean[2], TRUE ~ `PA(44:1)`),
   #      `PA(44:2)` = case_when(is.na(`PA(44:2)`) ~ df_mean_total$cal_mean[3], TRUE ~ `PA(44:2)`),
    #     `PG(34:1)` = case_when(is.na(`PG(34:1)`) ~ df_mean_total$cal_mean[4], TRUE ~ `PG(34:1)`),
     #    `TG(58:12)` = case_when(is.na(`TG(58:12)`) ~ df_mean_total$cal_mean[5], TRUE ~ `TG(58:12)`),
      #   `TG(42:2)` = case_when(is.na(`TG(42:2)`) ~ df_mean_total$cal_mean[6], TRUE ~ `TG(42:2)`),
       #  `TG(40:1)` = case_when(is.na(`TG(40:1)`) ~ df_mean_total$cal_mean[7], TRUE ~ `TG(40:1)`))
```


---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------


## Distribution
Normality check, prepare data frame for a longer format with the purpose of checking the distribution of data 
```{r echo=FALSE}
pivot_conc <- lip_df_imputed %>% 
  pivot_longer(cols = -c(Label, Sample_ID), 
               names_to = "Lipid", 
               values_to = "Concentrations")

# Calculate mean of every lipid across samples
mean_concentration <- pivot_conc %>% 
  group_by(Lipid) %>% 
  summarize(Mean = mean(Concentrations))

# Distribution of mean lipid concentration
mean_concentration %>% 
  ggplot(aes(x = Mean)) + 
  geom_density() + 
  labs(subtitle = "Distribution of lipid concentrations",
       x = "Lipid concentrations")
```

**Comments:** a skewed distribution, thus we need to normalize data.


---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------


## Distribution
Normality test, Kolmogorov-Smirnov Test
```{r echo=FALSE}
# Kolmogorov-Smirnov Test check if the concentrations are normaly distributed
pval_ks_test <- ks.test(x = mean_concentration$Mean, 
        y = "pnorm", 
        mean = mean(mean_concentration$Mean), 
        sd = sd(mean_concentration$Mean))$p.value
pval_ks_test
# Data is normal distributed if the p-value > 0.05. So the concentration variable is normally distributed.
```


---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------


# Transform/normalize/scale
Getting a dataframe with log and scaled concentrations
```{r echo=FALSE}
# Logtransform data
lipid_data_log <- lip_df_imputed
lipid_data_log[,3:ncol(lipid_data_log)] <- log2(lipid_data_log[,3:ncol(lipid_data_log)])

# Scale data 
lipid_data_scale <- lipid_data_log
lipid_data_scale[,3:ncol(lip_df_imputed)] <- scale(lipid_data_scale[,3:ncol(lipid_data_scale)])

# Save data frame 
lipid_data_log_scale <- as_tibble(lipid_data_scale)  %>% 
  write_csv(paste(path, "data/02_lipidomics_data_imputed_log2_zscore.csv", sep="/")) 
```


Distribution of log and scaled data
```{r echo=FALSE}
pivot_lipid_data <- lipid_data_scale %>% 
  pivot_longer(cols = -c(Sample_ID, Label), 
               names_to = "Lipid", 
               values_to = "Concentrations")

# Boxplot of raw metabolite concentrations
lipid_plot <- pivot_lipid_data %>% 
  ggplot(mapping = aes(x = Concentrations, y = Lipid)) + 
  geom_boxplot(alpha = 0.5) + 
  geom_point(position = position_jitter(seed = 0.4, width = 0.02), size = 0.8) +
  coord_flip() + 
  labs(x = 'Log2 and zscore lipid concentrations', title = "Distribution of lipid concentrations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.5))
lipid_plot
```


---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------


# Combine cleaned and log2/zscore data frames
```{r echo=FALSE}
combined_data <- left_join(lipid_data_log_scale, clinical_data_clean,  by = c("Label" = "visit", "Sample_ID" = "id")) %>% 
  select(Sample_ID, bwcat, randomization, Label, NAFLD, everything()) %>% 
  write_csv(paste(path, "data/02_combined_data_imputed_log2_zscore.csv", sep="/")) 
combined_data
```




#### Correlation: plot of log2 transformed and scaled metabolites
#Pearson's correlation is used (default)
```{r}
correlationplot_lipids <- plot_correlation(lipid_data_log_scale, type = "c")
#ggsave(paste(path, "results/02_01_metabolite_correlationplot.png", sep="/"), plot = correlationplot_met, device = "png", width = 10, height = 10)
```
**Comments:** Some correlations are identified between the metabolites

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------