---
title: "Lipidomics Baseline Analysis"
subtitle: "Comparion in the Systematical and Structural Composition of the Lipids"
author: "Sofie Olund Villumsen"
date: "2/9/2021"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

# Initial settings

```{r echo=FALSE}
# Clear workspace
rm(list = ls()) 
```

Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lipidomeR) 
```

Load data
```{r message=FALSE}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie/Baseline_study/Lipidomics"
setwd(path)

lipid_data <- read_csv(paste(path, "data/02_lipidomics_data_log2_zscore.csv", sep="/"))
combined_data <- read_csv(paste(path, "data/02_combined_data_log2_zscore.csv", sep="/"))

top_lipids <- read_tsv(paste(path, "data/04_lipidomics_top_lipids.tsv", sep="/"))
```

```{r echo=FALSE}
# Get a list of metabolites
lipid_col_names <- colnames(lipid_data)[3:ncol(lipid_data)]
```

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------


```{r echo=FALSE}
# Prepare data
lipid_data_baseline <- combined_data %>% 
  filter(Label == "A") %>% 
  unite("SampleID_bwcat_NAFLD", c(bwcat_NAFLD, Sample_ID), remove = FALSE) %>% 
  filter(SampleID_bwcat_NAFLD != "NBW_FALSE_050") %>% 
  select(-everything(), bwcat, bwcat_NAFLD, Sample_ID, Label, all_of(lipid_col_names))
```


# Two groups: NBW and LBW

```{r echo=FALSE}
lipid_data_baseline_x <- lipid_data_baseline %>% 
  select(bwcat, everything(), -c("bwcat_NAFLD","Sample_ID","Label"))

lip_names <- colnames(lipid_data_baseline_x[,2:ncol(lipid_data_baseline_x)])
lipid_mapping <- map_lipid_names( x = lip_names )

lipid_data_baseline_x$bwcat <- factor(lipid_data_baseline_x$bwcat, levels=c("LBW", "NBW"))

# Compute the regression models
result.limma <-
  compute_models_with_limma(
    x = lipid_data_baseline_x,
    dependent.variables = lipid_mapping$Name,
    independent.variables = c("bwcat"))
```

```{r echo=FALSE, warning=FALSE}
# Save as png file
png(paste(path, "results/05_01_lipidomeR_NBW_LBW_heatmaps.png", sep="/"),
    width = 6*300,        # 5 x 300 pixels
    height = 6*550,
    res = 200,            # 300 pixels per inch
    pointsize = 10)      # font size) 
figure.output <- heatmap_lipidome_from_limma(
     x = result.limma$"model",
     names.mapping = lipid_mapping,
     axis.x.carbons = FALSE,
     class.facet = "row",
     plot.all = TRUE,
     plot.individual = FALSE,
     print.figure = TRUE,
     scales = "free",
     space = "free"
 )
dev.off()
```

##### **Comments:** The plot shows that there is no significant difference between the abundance of lipids between the birthweight groups (NBW, LBW). There is a tendency of difference between the two groups in the abundance of (however not significantly):
- Phosphatidylcholines (PC)
- Phosphatidylethanolamine (PE)
- Sphingomyelin (SM) 
- Triglycerides (TG)

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

# Three groups: NBW, LBW÷NAFLD and LBW+NAFLD

```{r echo=FALSE}
lipid_data_baseline_y <- lipid_data_baseline %>% 
  mutate(Group = case_when(bwcat_NAFLD == "NBW_FALSE" ~ "NBW", 
                           bwcat_NAFLD == "LBW_FALSE" ~ "LBW÷NAFLD",
                           bwcat_NAFLD == "LBW_TRUE" ~ "LBW+NAFLD")) %>% 
  select(Group, everything(), -c("bwcat","bwcat_NAFLD","Sample_ID","Label"))

lip_names <- colnames(lipid_data_baseline_y[,2:ncol(lipid_data_baseline_y)])
lipid_mapping <- map_lipid_names( x = lip_names )

# Create factors for the order
lipid_data_baseline_y$Group <- factor(lipid_data_baseline_y$Group, 
                                                   levels=c("NBW","LBW÷NAFLD","LBW+NAFLD"))

# Compute the regression models
result.limma <-
  compute_models_with_limma(
    x = lipid_data_baseline_y,
    dependent.variables = lipid_mapping$Name,
    independent.variables = c("Group"))

```


```{r echo=FALSE, warning=FALSE}
# Save as png file
png(paste(path, "results/05_02_lipidomeR_NBW_LBWnoNAFLD_LBWNAFLD_heatmaps.png", sep="/"),
    width = 6*300,        # 5 x 300 pixels
    height = 6*550,
    res = 200,            # 300 pixels per inch
    pointsize = 10)      # font size) 
figure.output <- heatmap_lipidome_from_limma(
     x = result.limma$"model",
     names.mapping = lipid_mapping,
     axis.x.carbons = FALSE,
     class.facet = "row",
     plot.all = TRUE,
     plot.individual = FALSE,
     print.figure = TRUE,
     scales = "free",
     space = "free"
 )
dev.off()
```

##### **Comments:** The plot shows that there is no significant difference between the abundance of lipids between the group LBW÷NAFLD and NBW. However, when comparing LBW÷NAFLD with LBW+NAFLD we significant difference in the abundance between the groups. The plot shows a differential abundance of (FDR < 0.1): 
- Increased: Ceramide (CER) [with unsaturated long-chain-fatty-acids (LCFA)]
- Increased: Lysophosphatidylcholines (LPC) [with saturated LCFA]
- Increased: Phosphatidylcholines (PC) [with saturated/unsaturated LCFA]
- Increased: Phosphatidylethanolamine (PE) [with unsaturated LCFA]
- Increased: Sphingomyelin (SM) [with unsaturated very-LCFA]
- Increased: Triglycerides (TG) [with saturated/unsaturated LCFA/VLCFA]
- Decreased: Cholesteaol Esters (CE) [with unsaturated LCFA]
- Decreased: Phosphatidylglycerol (PG) [with unsaturated LCFA]
- Decreased: Triglycerides (TG) [with unsaturated VLCFA]

Furthermore, we observe that when the LBW+NAFLD is removed, the difference bewteen NBW and LBW÷NAFLD is decreasing. 

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

# Top lipids 
Three groups: NBW, LBW÷NAFLD and LBW+NAFLD

```{r echo=FALSE}
top_lipids_list <- unique(top_lipids$Lipids)

lipid_data_baseline_z <- lipid_data_baseline %>% 
  mutate(Group = case_when(bwcat_NAFLD == "NBW_FALSE" ~ "NBW", 
                           bwcat_NAFLD == "LBW_FALSE" ~ " LBW w/o NAFLD",
                           bwcat_NAFLD == "LBW_TRUE" ~ " LBW w/ NAFLD")) %>% 
  select(Group, all_of(top_lipids_list))

lip_names <- colnames(lipid_data_baseline_z[,2:ncol(lipid_data_baseline_z)])
lipid_mapping <- map_lipid_names( x = lip_names )

# Create factors for the order
lipid_data_baseline_z$Group <- factor(lipid_data_baseline_z$Group, 
                                                   levels=c("NBW"," LBW w/o NAFLD"," LBW w/ NAFLD"))

# Compute the regression models
result.limma <-
  compute_models_with_limma(
    x = lipid_data_baseline_z,
    dependent.variables = lipid_mapping$Name,
    independent.variables = c("Group"))

```


```{r echo=FALSE, warning=FALSE}
# Save as png file
png(paste(path, "results/05_03_lipidomeR_toplipids_NBW_LBWnoNAFLD_LBWNAFLD_heatmaps.png", sep="/"),
    width = 6*300,        # 5 x 300 pixels
    height = 6*250,
    res = 200,            # 300 pixels per inch
    pointsize = 10)      # font size) 
figure.output <- heatmap_lipidome_from_limma(
     x = result.limma$"model",
     names.mapping = lipid_mapping,
     axis.x.carbons = FALSE,
     class.facet = "row",
     plot.all = TRUE,
     plot.individual = FALSE,
     print.figure = TRUE,
     scales = "free",
     space = "free"
 )
dev.off()
```

