---
title: "Lipidomics Baseline Analysis"
subtitle: "Biological Network Topology Analysis"
author: "Sofie Olund Villumsen"
date: "30/9/2021"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

# Initial settings
```{r echo=FALSE}
# Clear workspace
rm(list = ls()) 
```

Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(correlation) # Compute rho and p-values of the correlation test for several pairs of variables at once
library(igraph)
library(leidenAlg) # Community detection
```

Load data
```{r message=FALSE}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie/Baseline_study/Lipidomics"
setwd(path)

lipid_data <- read_csv(paste(path, "data/02_lipidomics_data_imputed_log2_zscore.csv", sep="/"))
lip_clin_data <- read_csv(paste(path, "data/02_combined_data_imputed_log2_zscore.csv", sep="/"))
top_lipids <- read_tsv(paste(path, "data/04_lipidomics_top_lipids.tsv", sep="/"))
```


Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
```{r}
set.seed(422053)
```


```{r echo=FALSE}
# Get a list of lipids
lip_col_names <- colnames(lipid_data)[3:ncol(lipid_data)]
```

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------


# Data preparation
```{r}
lip_df <- lipid_data %>% 
  filter(Label == "A") 

lip_df_2 <- lip_df %>% 
  select(-everything(), all_of(lip_col_names)) %>% 
  t()

colnames(lip_df_2) <- lip_df$Sample_ID
rownames(lip_df_2) <- lip_col_names
lip_df_2 <- as_tibble(lip_df_2) %>% 
  mutate(lipids = lip_col_names) %>% 
  select(lipids, everything()) 

##write_csv(kegg, paste(path, "data/05_pathway.csv", sep="/")) 
```


------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------


# Correlation Analysis 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.4

lipid_conc <- lip_df %>% 
  select(everything(), -c(Label, Sample_ID))

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix <- correlation(lipid_conc, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```



------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------


# Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
corr_matrix_all <- corr_matrix %>%
  rename(Lipid1 = Parameter1, 
         Lipid2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Lipid1, Lipid2, rho) %>% 
  mutate(Lipid1 = gsub(",", "_", Lipid1),
         Lipid2 = gsub(",", "_", Lipid2)) %>% 
  as.data.frame()

write_csv(corr_matrix_all, paste(path, "data/06_posneg_edge_table.csv", sep="/"))

# Graph of associations, weighted
g_all <- graph_from_data_frame(corr_matrix_all, directed = FALSE)
g_all <- set_edge_attr(g_all, "weight", value= corr_matrix_all$rho)
is_weighted(g_all)

# Plot graph
plot(g_all, vertex.label = V(g_all)$name)
```



## Random network 
```{r}
# Number of nodes in positive network
n_nodes <- unique(c(corr_matrix_all$Lipid1, corr_matrix_all$Lipid2))

# Graph of random network, unweighted, node and edge number based on a network of the same size
g_random <- erdos.renyi.game(n = length(n_nodes),           # Number of nodes in the graph
                             p.or.m = length(corr_matrix_all$rho), # Number of edges in the graph
                             type = "gnm",
                             directed = FALSE)

# Leidenalg: Community detection algorithm 
leiden_random <- leiden.community(g_random)

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_random, membership(leiden_random))

degree_distribution(g_random)
plot(g_random)
```


## Comparison: pos+neg network
vs random network
```{r}
all_char <- c(gorder(g_all), 
              gsize(g_all),
              edge_density(g_all),
              transitivity(g_all, type = ("undirected")), # global clustering coefficient
              min_cut(g_all)
              )

random_char <- c(gorder(g_random), 
                 gsize(g_random), 
                 edge_density(g_random),
                 transitivity(g_random, type = ("undirected")), # global clustering coefficient
                 min_cut(g_random)
                 )

network_characteristics <- as_tibble(rbind(all_char, random_char)) %>% 
  mutate(network_type = c("pos_neg_network", "random_network")) %>% 
  rename(n_nodes = "V1", n_edges = "V2", density = "V3", cluster_coeff = "V4", min_cut ="V5") %>% 
  select(network_type, everything())
network_characteristics
```

Density = the number of edges a node has, divided by the total possible edges a node could have
....


## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
# Leidenalg: Community detection algorithm 
leiden_community <- leiden.community(g_all)

# Community seperation
g_all_community <- as_tibble(leiden_community$membership)  %>% 
  mutate(Lipids = leiden_community$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_all, membership(leiden_community))

# Number of communities and how many metabolites each contain
sizes(leiden_community)


##wc <- cluster_walktrap(g_pos)
##plot(wc, g_pos)

##plot(leiden_community$names, g_pos)
```

```{r}
# Create table which contain information on the metabolites
node_table_all <- g_all_community %>% 
  rename(Community_nr = value) %>% 
  select(Lipids, Community_nr)

node_table_all2 <- left_join(node_table_all, fdr_lipids, by = c("Lipids" = "Lipids"), keep = TRUE) %>% 
  rename(Lipids = Lipids.x, fdr_lipids = Lipids.y)

node_table_all2
write_csv(node_table_all2, paste(path, "data/06_posneg_node_table.csv", sep="/"))
```

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------



# Positive correlation network
```{r}
corr_matrix_pos <- corr_matrix_df %>% 
  filter(rho > 0) %>% 
  select(Lipid1, Lipid2, rho) %>% 
  mutate(Lipid1 = gsub(",", "_", Lipid1),
         Lipid2 = gsub(",", "_", Lipid2)) %>% 
  as.data.frame()

write_csv(corr_matrix_pos, paste(path, "data/06_pos_edge_table.csv", sep="/"))

# Graph of positive associations, weighted
g_pos <- graph_from_data_frame(corr_matrix_pos, directed = FALSE)
g_pos <- set_edge_attr(g_pos, "weight", value= corr_matrix_pos$rho)
is_weighted(g_pos)

# Plot graph
plot(g_pos, vertex.label = V(g_pos)$name)
```
The above plot shows a biological network with weights (spearmann). A network makes it easier to discover trends/groupings in data and give us a more holistic view of the correlations between the molecules. By applying a community analysis (leiden algorithm) we are able to discover biochemicals with common properties and keep them tightly connected (check among others the modularity score, as a step in the robustness check of the network). 


## Random network
```{r}
# Number of nodes in positive network
n_nodes <- unique(c(corr_matrix_pos$Lipid1, corr_matrix_pos$Lipid2))

# Graph of random network, unweighted, node and edge number based on a network of the same size
g_random <- erdos.renyi.game(n = length(n_nodes),           # Number of nodes in the graph
                             p.or.m = length(corr_matrix_pos$rho), # Number of edges in the graph
                             type = "gnm",
                             directed = FALSE)
degree_distribution(g_random)
plot(g_random)
```


## Comparison: positive vs random network
```{r}
pos_char <- c(gorder(g_pos), 
              gsize(g_pos),
              diameter(g_pos, directed = FALSE, unconnected = TRUE, weights = NULL), 
              edge_density(g_pos),
              transitivity(g_pos, type = ("undirected")), # global clustering coefficient
              min_cut(g_pos)
              )

random_char <- c(gorder(g_random), 
                 gsize(g_random), 
                 diameter(g_random, directed = FALSE, unconnected = TRUE, weights = NULL), 
                 edge_density(g_random),
                 transitivity(g_random, type = ("undirected")), # global clustering coefficient
                 min_cut(g_random)
                 )

network_characteristics <- as_tibble(rbind(pos_char, random_char)) %>% 
  mutate(network_type = c("pos_network", "random_network")) %>% 
  rename(n_nodes = "V1", n_edges = "V2", diameter = "V3", density = "V4", cluster_coeff = "V5", min_cut ="V6") %>% 
  select(network_type, everything())
network_characteristics
```

Diameter = a higher diameter indicates that there are some nodes in the perifery (classic for a biological network)
Density = the number of edges a node has, divided by the total possible edges a node could have
....


## Positive graph: Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
# Leidenalg: Community detection algorithm 
leiden_community <- leiden.community(g_pos)

# Community seperation
g_pos_community <- as_tibble(leiden_community$membership)  %>% 
  mutate(Lipids = leiden_community$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_pos, membership(leiden_community))

# Number of communities and how many metabolites each contain
sizes(leiden_community)

##wc <- cluster_walktrap(g_pos)
##plot(wc, g_pos)

##plot(leiden_community$names, g_pos)
```


```{r}
# Create table which contain information on the metabolites
node_table_pos <- g_pos_community %>% 
  rename(Community_nr = value) %>% 
  select(Lipids, Community_nr)

node_table_pos2 <- left_join(node_table_pos, fdr_lipids, by = c("Lipids" = "Lipids"), keep = TRUE) %>% 
  rename(Lipids = Lipids.x, fdr_lipids = Lipids.y)

node_table_pos2
write_csv(node_table_pos2, paste(path, "data/06_pos_node_table.csv", sep="/"))
```


# Random graph
```{r}
# Leidenalg: Community detection algorithm 
leiden_random <- leiden.community(g_random)

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_random, membership(leiden_random))
```


- https://igraph.org/r/doc/ 




-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------

# NBW - lipidomics
```{r}
nbw <- lip_clin_data %>% 
  filter(Label == "A", bwcat == "NBW") %>% 
  select(-everything(), all_of(lip_col_names))

limma_LBWvsNBW <- read_csv(paste(path, "data/04_lipidomics_limma2_LBWvsNBW_imputed.csv", sep="/"), show_col_types = FALSE)
```

## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.35

biochem_conc_nbw <- nbw %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_nbw <- correlation(biochem_conc_nbw, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
corr_matrix_nbw_posneg <- corr_matrix_nbw %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_nbw_posneg, paste(path, "data/03_NBW_baseline_lipidomics_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_nbw <- graph_from_data_frame(corr_matrix_nbw_posneg, directed = FALSE)
g_posneg_nbw <- set_edge_attr(g_posneg_nbw, "weight", value= corr_matrix_nbw_posneg$rho)
is_weighted(g_posneg_nbw)

# Plot graph
plot(g_posneg_nbw, vertex.label = V(g_posneg_nbw)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_nbw <- leiden.community(g_posneg_nbw)

# Community seperation
g_posneg_nbw_community <- as_tibble(leiden_community_nbw$membership)  %>% 
  mutate(Biochemicals = leiden_community_nbw$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_nbw, membership(leiden_community_nbw))

# Number of communities and how many metabolites each contain
sizes(leiden_community_nbw)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_nbw_posneg <- g_posneg_nbw_community %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr)

pvalue_lipids_NBW <- top_lipids %>% 
  select(Lipids)

node_table_nbw_posneg2 <- left_join(node_table_nbw_posneg, limma_LBWvsNBW, by = c("Biochemicals" = "Lipids"), keep = TRUE) 
node_table_nbw_posneg3 <- left_join(node_table_nbw_posneg2, pvalue_lipids_NBW, by = c("Biochemicals" = "Lipids"), keep = TRUE) %>% rename(pvalue_metabolites = Lipids.y, Lipids = Lipids.x) %>% 
  select(Lipids, Community_nr, logFC, P.Value, adj.P.Val, pvalue_metabolites, HMDB, KEGG, PubChem )

node_table_nbw_posneg3
write_csv(node_table_nbw_posneg3, paste(path, "data/03_NBW_baseline_lipidomics_posneg_node_table_imputed.csv", sep="/"))
```



-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------


# LBW - lipidomics
```{r}
lbw <- lip_clin_data %>% 
  filter(Label == "A", bwcat == "LBW") %>% 
  select(-everything(), all_of(lip_col_names))

limma_LBWvsNBW <- read_csv(paste(path, "data/04_lipidomics_limma2_LBWvsNBW_imputed.csv", sep="/"), show_col_types = FALSE)
```

## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.35

biochem_conc_lbw <- lbw %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_lbw <- correlation(biochem_conc_lbw, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
corr_matrix_lbw_posneg <- corr_matrix_lbw %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_lbw_posneg, paste(path, "data/03_LBW_baseline_lipidomics_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_lbw <- graph_from_data_frame(corr_matrix_lbw_posneg, directed = FALSE)
g_posneg_lbw <- set_edge_attr(g_posneg_lbw, "weight", value= corr_matrix_lbw_posneg$rho)
is_weighted(g_posneg_lbw)

# Plot graph
plot(g_posneg_lbw, vertex.label = V(g_posneg_lbw)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_lbw <- leiden.community(g_posneg_lbw)

# Community seperation
g_posneg_lbw_community <- as_tibble(leiden_community_lbw$membership)  %>% 
  mutate(Biochemicals = leiden_community_lbw$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_lbw, membership(leiden_community_lbw))

# Number of communities and how many metabolites each contain
sizes(leiden_community_lbw)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_lbw_posneg <- g_posneg_lbw_community %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr)

pvalue_lipids_LBW <- top_lipids %>% 
  select(Lipids)

node_table_lbw_posneg2 <- left_join(node_table_lbw_posneg, limma_LBWvsNBW, by = c("Biochemicals" = "Lipids"), keep = TRUE) 
node_table_lbw_posneg3 <- left_join(node_table_lbw_posneg2, pvalue_lipids_LBW, by = c("Biochemicals" = "Lipids"), keep = TRUE) %>% rename(pvalue_lipids = Lipids.y, Lipids = Lipids.x) %>% 
  select(Lipids, Community_nr, logFC, P.Value, adj.P.Val, pvalue_lipids, HMDB, KEGG, PubChem )

node_table_lbw_posneg3
write_csv(node_table_lbw_posneg3, paste(path, "data/03_LBW_baseline_lipidomics_posneg_node_table_imputed.csv", sep="/"))
```


-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------

# -- OMICS NETWORKS

```{r}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie/Baseline_study/Metabolomics"
setwd(path)

metabo_data <- read_csv(paste(path, "data/02_combined_data_imputed_log2_zscore.csv", sep="/"), show_col_types = FALSE)
top_metabolites <- read_tsv(paste(path, "data/04_metabolomics_top_metabolites.tsv", sep="/"), show_col_types = FALSE)

top_met_df <- metabo_data %>% 
  unite("SampleID_bwcat_NAFLD", c(bwcat, NAFLD, Sample_ID), remove = FALSE) %>% 
  filter(Label == "A", bwcat == "NBW") %>% 
  select(SampleID_bwcat_NAFLD, all_of(top_metabolites$Metabolites))

path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie/Baseline_study/Lipidomics"
setwd(path)
```



# NBW: omics
```{r}
nbw_lip_df <- lip_clin_data %>% 
  unite("SampleID_bwcat_NAFLD", c(bwcat, NAFLD, Sample_ID), remove = FALSE) %>%
  filter(Label == "A", bwcat == "NBW") %>% 
  select(-everything(), SampleID_bwcat_NAFLD, all_of(lip_col_names))

nbw_lip_topmet_df <- left_join(nbw_lip_df, top_met_df, by = c("SampleID_bwcat_NAFLD" = "SampleID_bwcat_NAFLD"), keep = FALSE) %>% 
  select(everything(), -SampleID_bwcat_NAFLD)
```

## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

biochem_conc_nbw_lipmet <- nbw_lip_topmet_df %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_nbw_lipmet <- correlation(biochem_conc_nbw_lipmet, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.40

corr_matrix_nbw_posneg_lipmet <- corr_matrix_nbw_lipmet %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_nbw_posneg_lipmet, paste(path, "data/03_NBW_baseline_lipidomics_topmet_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_nbw_lipmet <- graph_from_data_frame(corr_matrix_nbw_posneg_lipmet, directed = FALSE)
g_posneg_nbw_lipmet <- set_edge_attr(g_posneg_nbw_lipmet, "weight", value= corr_matrix_nbw_posneg_lipmet$rho)
is_weighted(g_posneg_nbw_lipmet)

# Plot graph
plot(g_posneg_nbw_lipmet, vertex.label = V(g_posneg_nbw_lipmet)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_nbw_lipmet <- leiden.community(g_posneg_nbw_lipmet)

# Community seperation
g_posneg_nbw_community_lipmet <- as_tibble(leiden_community_nbw_lipmet$membership)  %>% 
  mutate(Biochemicals = leiden_community_nbw_lipmet$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_nbw_lipmet, membership(leiden_community_nbw_lipmet))

# Number of communities and how many metabolites each contain
sizes(leiden_community_nbw_lipmet)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_nbw_lipmet_posneg <- g_posneg_nbw_community_lipmet %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr)

pvalue_lipids_NBW_lipmet <- top_lipids %>%
  filter(group1 == "LBW-NAFLD") %>% 
  select(Lipids) %>% unique()

pvalue_met_NBW_lipmet <- top_metabolites %>% 
  filter(group1 == "LBW-NAFLD") %>% 
  select(Metabolites) %>%  unique() %>% rename(Lipids = Metabolites)

pvalue_metlip <- rbind(pvalue_lipids_NBW_lipmet, pvalue_met_NBW_lipmet)

node_table_nbw_lipmet_posneg2 <- left_join(node_table_nbw_lipmet_posneg, pvalue_metlip, by = c("Biochemicals" = "Lipids"), keep = TRUE) %>% 
  rename(pvalue_biomolecules = Lipids)

node_table_nbw_lipmet_posneg2
write_csv(node_table_nbw_lipmet_posneg2, paste(path, "data/03_NBW_baseline_lipidomics_topmet_posneg_node_table_imputed.csv", sep="/"), na = "")
```


-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------


# LBW: omics
```{r}
lbw_lip_df <- lip_clin_data %>% 
  unite("SampleID_bwcat_NAFLD", c(bwcat, NAFLD, Sample_ID), remove = FALSE) %>%
  filter(Label == "A", bwcat == "LBW") %>% 
  select(-everything(), SampleID_bwcat_NAFLD, all_of(lip_col_names))

lbw_lip_topmet_df <- left_join(lbw_lip_df, top_met_df, by = c("SampleID_bwcat_NAFLD" = "SampleID_bwcat_NAFLD"), keep = FALSE) %>% 
  select(everything(), -SampleID_bwcat_NAFLD)
```


## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

biochem_conc_lbw_lipmet <- lbw_lip_topmet_df %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_lbw_lipmet <- correlation(biochem_conc_lbw_lipmet, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.40

corr_matrix_lbw_lipmet_posneg <- corr_matrix_lbw_lipmet %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_lbw_lipmet_posneg, paste(path, "data/03_LBW_baseline_lipidomics_topmet_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_lbw_lipmet <- graph_from_data_frame(corr_matrix_lbw_lipmet_posneg, directed = FALSE)
g_posneg_lbw_lipmet <- set_edge_attr(g_posneg_lbw_lipmet, "weight", value= corr_matrix_lbw_lipmet_posneg$rho)
is_weighted(g_posneg_lbw_lipmet)

# Plot graph
plot(g_posneg_lbw_lipmet, vertex.label = V(g_posneg_lbw_lipmet)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_lbw_lipmet <- leiden.community(g_posneg_lbw_lipmet)

# Community seperation
g_posneg_lbw_community_lipmet <- as_tibble(leiden_community_lbw_lipmet$membership)  %>% 
  mutate(Biochemicals = leiden_community_lbw_lipmet$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_lbw_lipmet, membership(leiden_community_lbw_lipmet))

# Number of communities and how many metabolites each contain
sizes(leiden_community_lbw_lipmet)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_lbw_posneg_lipmet <- g_posneg_lbw_community_lipmet %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr)

node_table_lbw_posneg_lipmet2 <- left_join(node_table_lbw_posneg_lipmet, pvalue_metlip, by = c("Biochemicals" = "Lipids"), keep = TRUE) %>% 
  rename(pvalue_biomolecules = Lipids)

node_table_lbw_posneg_lipmet2
write_csv(node_table_lbw_posneg_lipmet2, paste(path, "data/03_LBW_baseline_lipidomics_topmet_posneg_node_table_imputed.csv", sep="/"), na="")
```



-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------


# LBW-NAFLD: omics
```{r}
lbw_noNAFLD_lip_df <- lip_clin_data %>% 
  unite("SampleID_bwcat_NAFLD", c(bwcat, NAFLD, Sample_ID), remove = FALSE) %>%
  filter(Label == "A", bwcat == "LBW", NAFLD == "FALSE") %>% 
  select(-everything(), SampleID_bwcat_NAFLD, all_of(lip_col_names))

lbw_noNAFLD_lip_topmet_df <- left_join(lbw_noNAFLD_lip_df, top_met_df, by = c("SampleID_bwcat_NAFLD" = "SampleID_bwcat_NAFLD"), keep = FALSE) %>% 
  select(everything(), -SampleID_bwcat_NAFLD)
```


## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

biochem_conc_lbw_noNAFLD_lipmet <- lbw_noNAFLD_lip_topmet_df %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_lbw_noNAFLD_lipmet <- correlation(biochem_conc_lbw_noNAFLD_lipmet, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.40

corr_matrix_lbw_noNAFLD_lipmet_posneg <- corr_matrix_lbw_noNAFLD_lipmet %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_lbw_noNAFLD_lipmet_posneg, paste(path, "data/03_LBW_noNAFLD_baseline_lipidomics_topmet_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_lbw_noNAFLD_lipmet <- graph_from_data_frame(corr_matrix_lbw_noNAFLD_lipmet_posneg, directed = FALSE)
g_posneg_lbw_noNAFLD_lipmet <- set_edge_attr(g_posneg_lbw_noNAFLD_lipmet, "weight", value= corr_matrix_lbw_noNAFLD_lipmet_posneg$rho)
is_weighted(g_posneg_lbw_noNAFLD_lipmet)

# Plot graph
plot(g_posneg_lbw_noNAFLD_lipmet, vertex.label = V(g_posneg_lbw_noNAFLD_lipmet)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_lbw_noNAFLD_lipmet <- leiden.community(g_posneg_lbw_noNAFLD_lipmet)

# Community seperation
g_posneg_lbw_noNAFLD_community_lipmet <- as_tibble(leiden_community_lbw_noNAFLD_lipmet$membership)  %>% 
  mutate(Biochemicals = leiden_community_lbw_noNAFLD_lipmet$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_lbw_noNAFLD_lipmet, membership(leiden_community_lbw_noNAFLD_lipmet))

# Number of communities and how many metabolites each contain
sizes(leiden_community_lbw_noNAFLD_lipmet)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_lbw_noNAFLD_posneg_lipmet <- g_posneg_lbw_noNAFLD_community_lipmet %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr)

node_table_lbw_noNAFLD_posneg_lipmet2 <- left_join(node_table_lbw_noNAFLD_posneg_lipmet, pvalue_metlip, by = c("Biochemicals" = "Lipids"), keep = TRUE) %>% 
  rename(pvalue_biomolecules = Lipids)

node_table_lbw_noNAFLD_posneg_lipmet2
write_csv(node_table_lbw_noNAFLD_posneg_lipmet2, paste(path, "data/03_LBW_noNAFLD_baseline_lipidomics_topmet_posneg_node_table_imputed.csv", sep="/"), na="")
```



-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------


# LBW+NAFLD: omics
```{r}
lbw_NAFLD_lip_df <- lip_clin_data %>% 
  unite("SampleID_bwcat_NAFLD", c(bwcat, NAFLD, Sample_ID), remove = FALSE) %>%
  filter(Label == "A", bwcat == "LBW", NAFLD == "TRUE") %>% 
  select(-everything(), SampleID_bwcat_NAFLD, all_of(lip_col_names))

lbw_NAFLD_lip_topmet_df <- left_join(lbw_NAFLD_lip_df, top_met_df, by = c("SampleID_bwcat_NAFLD" = "SampleID_bwcat_NAFLD"), keep = FALSE) %>% 
  select(everything(), -SampleID_bwcat_NAFLD)
```


## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

biochem_conc_lbw_NAFLD_lipmet <- lbw_NAFLD_lip_topmet_df %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_lbw_NAFLD_lipmet <- correlation(biochem_conc_lbw_NAFLD_lipmet, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.40

corr_matrix_lbw_NAFLD_lipmet_posneg <- corr_matrix_lbw_NAFLD_lipmet %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_lbw_NAFLD_lipmet_posneg, paste(path, "data/03_LBW_NAFLD_baseline_lipidomics_topmet_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_lbw_NAFLD_lipmet <- graph_from_data_frame(corr_matrix_lbw_NAFLD_lipmet_posneg, directed = FALSE)
g_posneg_lbw_NAFLD_lipmet <- set_edge_attr(g_posneg_lbw_NAFLD_lipmet, "weight", value= corr_matrix_lbw_NAFLD_lipmet_posneg$rho)
is_weighted(g_posneg_lbw_NAFLD_lipmet)

# Plot graph
plot(g_posneg_lbw_noNAFLD_lipmet, vertex.label = V(g_posneg_lbw_noNAFLD_lipmet)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_lbw_NAFLD_lipmet <- leiden.community(g_posneg_lbw_NAFLD_lipmet)

# Community seperation
g_posneg_lbw_NAFLD_community_lipmet <- as_tibble(leiden_community_lbw_NAFLD_lipmet$membership)  %>% 
  mutate(Biochemicals = leiden_community_lbw_NAFLD_lipmet$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_lbw_NAFLD_lipmet, membership(leiden_community_lbw_NAFLD_lipmet))

# Number of communities and how many metabolites each contain
sizes(leiden_community_lbw_NAFLD_lipmet)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_lbw_NAFLD_posneg_lipmet <- g_posneg_lbw_NAFLD_community_lipmet %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr)

node_table_lbw_NAFLD_posneg_lipmet2 <- left_join(node_table_lbw_NAFLD_posneg_lipmet, pvalue_metlip, by = c("Biochemicals" = "Lipids"), keep = TRUE) %>% 
  rename(pvalue_biomolecules = Lipids)

node_table_lbw_NAFLD_posneg_lipmet2
write_csv(node_table_lbw_NAFLD_posneg_lipmet2, paste(path, "data/03_LBW_NAFLD_baseline_lipidomics_topmet_posneg_node_table_imputed.csv", sep="/"))
```






