---
title: "Metabolomics Baseline Analysis"
subtitle: "Data preprocessing"
author: "Sofie Olund Villumsen"
date: "23/8/2021"
output: 
  html_document:
    toc: true
    toc_depth: 2
---
R markdown doc analyzing and exploring metabolomics data in combination with clinical from the type 2 diabetes overfeeding study of 48 men with mean age of 38 years.

# Initial settings

Clear workspace
```{r echo=FALSE}
rm(list = ls()) 
```

Load packages
```{r message=FALSE}
library(tidyverse)
library(caret) # Function nearzerovar
library(impute) # Imputing data with KNN
library(DataExplorer) # http://boxuancui.github.io/DataExplorer/
```

Load data
```{r message=FALSE}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie/Baseline_study/Metabolomics"
setwd(path)

metabo_data <- read_csv(paste(path, "data/01_metabolomics_data.csv", sep="/")) 
clinical_data <- read_csv(paste(path, "data/01_clinical_data.csv", sep="/"))
```


---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------


# Tidy and combine metabolomics and clinical data

```{r echo=FALSE}
# Metabolomics data cleanup
# Filter data to contain samples from the baseline only
metabolite_data_clean <- metabo_data %>% 
  select(...3, ...2, everything(), -...1) %>% 
  slice(8:140) %>% 
  rename(Sample_ID = ...2, Label = ...3) %>% 
  mutate(Label = str_to_upper(Label), 
         Sample_ID = case_when(nchar(Sample_ID) == 1 ~ as.character(paste0("00", Sample_ID)), 
                               nchar(Sample_ID) == 2 ~ as.character(paste0("0", Sample_ID))))
```

```{r echo=FALSE}
# Clinical data cleanup
clinical_data_clean <- clinical_data %>% 
  unite("bwcat_NAFLD", c(bwcat,NAFLD), remove = FALSE) %>% 
  filter(visit != "M") %>% 
  filter(visitid != "C001")
```

Combine cleaned data frames
```{r echo=FALSE}
combined_data <- left_join(metabolite_data_clean, clinical_data_clean,  by = c("Label" = "visit", "Sample_ID" = "id")) %>% 
  select(Sample_ID, bwcat, randomization, Label, NAFLD, everything()) #%>% 
 # write_csv(paste(path, "data/07_combined_data.csv", sep="/")) 
combined_data
```

Database information
```{r echo=FALSE}
# Database
metabolite_database <- metabo_data %>% 
  select(everything(), -c(...1,...2)) %>% 
  slice(1:4) %>% 
  rename(database_met = ...3) %>% 
  t()

colnames(metabolite_database) <- c("HMDB",  "KEGG" ,  "PubChem" , "CheBI") #colnames(metabo_data)[4:ncol(metabo_data)]
metabolite_database <- as_tibble(metabolite_database) %>% 
 filter(KEGG != "KEGG") %>% 
  mutate(metabolites = colnames(metabo_data)[4:ncol(metabo_data)])

as.data.frame(metabolite_database)  %>% 
  write_csv(paste(path, "data/02_metabolite_database_info.csv", sep="/"))
```


---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------



# Explore data

## Variance check (remove cols with zero or near zero values)
```{r echo=FALSE}
nearZeroVar(metabolite_data_clean[, 3:ncol(metabolite_data_clean)], saveMetrics = TRUE) %>% 
  filter(nzv == TRUE)
```
**Comments:** No variance data, thus no need for removing any variables.

```{r echo=FALSE}
# Metabolomics data filter
metabolite_data <- combined_data %>% 
  select(all_of(colnames(metabolite_data_clean))) %>% 
  mutate(across(.cols = 2:ncol(metabolite_data_clean), as.double))
```


## Missing values in combined data
```{r echo=FALSE}
profile_missing(metabolite_data_clean) %>% arrange(desc(pct_missing))
```
**Comments:** 1 missing value in the data set (alpha-ketoglutaric acid)

## Identify negative values in metaboloics data
```{r echo=FALSE}
minvals <- as.data.frame(apply(metabolite_data_clean,2,min))
minvals
```
**Comments:** From the list of min vals, we see that the minimum values for each metabolite differs a lot and some metabolites have negative concentration. Missing values and negative concentrations will be imputed. Afterwards will esch metabolite be normalized and then scaled.

## Outlier overview via boxplot
```{r echo=FALSE}
# Pivot data frame 
pivot_metabo_data <- metabolite_data %>% 
  pivot_longer(cols = -c(Sample_ID, Label), 
               names_to = "Metabolite", 
               values_to = "Concentrations")

# Boxplot of raw metabolite concentrations
metabolite_plot <- pivot_metabo_data %>% 
  ggplot(mapping = aes(x = Concentrations, y = Metabolite)) + 
  geom_boxplot(alpha = 0.5) + 
  geom_point(position = position_jitter(seed = 0.4, width = 0.02), size = 0.8) +
  coord_flip() + 
  labs(x = 'Raw metabolite concentrations', title = "Distribution of metabolite concentrations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.5))
metabolite_plot
```

**Comments:** from this plot it is clear that the metabolites follow different ranges and that outliers do occur.The following metabolites will be further explored: 


## Outliers to NAN
```{r}
met_names <- metabolite_data %>% 
  select(everything(), -c(Sample_ID, Label))

met_df <- metabolite_data %>% 
  unite("Label_SampleID", c(Label,Sample_ID), remove = TRUE) 

df_rm_outlier <- data.frame(met_df$Label_SampleID)

# Loop for identifying outliers and replacing them with NA values
for (met in colnames(met_names)) {
  # Identify outliers
  x_new <- filter(met_df, met_df[[met]] %in% boxplot(met_df[[met]], plot = FALSE)$out)
  
  # Replace the outlier values with NA
  metabolite <- replace(met_df[[met]], met_df[[met]] %in% x_new[[met]], values = NA)
  
  # Combine in data frame
  df_rm_outlier <- cbind(df_rm_outlier, data.frame(metabolite))
}

colnames(df_rm_outlier) <- colnames(met_df)
```

How many Nan values is there in the data set
```{r}
table(is.na(df_rm_outlier))
sapply(df_rm_outlier, function(x) sum(is.na(x)))
```


## Impute data with k-nearest neighbor
```{r}
# Convert data frame from numeric to factor
indx <- sapply(df_rm_outlier, is.numeric)
df_rm_outlier[indx] <- lapply(df_rm_outlier[indx], function(df_rm_outlier) as.factor(as.character(df_rm_outlier)))
df_rm_outlier <- df_rm_outlier[2:66]

# Set random seed 
if(exists(".Random.seed")) rm(.Random.seed)

# Impute with KNN function
df_rm_outlier_imputed <- impute.knn(as.matrix(df_rm_outlier))

met_df_imputed <- as_tibble(df_rm_outlier_imputed$data) %>% 
  mutate(Sample_ID = metabolite_data$Sample_ID,
         Label = metabolite_data$Label) %>% 
  select(Sample_ID, Label, everything())
```





---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------



# Distribution
Normality check, prepare data frame for a longer format with the purpose of checking the distribution of data 
```{r echo=FALSE}
pivot_conc <- met_df_imputed %>% 
  pivot_longer(cols = -c(Label, Sample_ID), 
               names_to = "Metabolite", 
               values_to = "Concentrations")

# Calculate mean of every metabolite across samples
mean_concentration <- pivot_conc %>% 
  group_by(Metabolite) %>% 
  summarize(Mean = mean(Concentrations))

# Distribution of mean metabolite concentration
mean_concentration %>% 
  ggplot(aes(x = Mean)) + 
  geom_density() + 
  labs(subtitle = "Distribution of metabolite concentrations",
       x = "Metabolite concentrations")
```
**Comments:** a skewed distribution, thus we need to normalize data.


# Distribution
Normality test, Kolmogorov-Smirnov Test
```{r echo=FALSE}
# Kolmogorov-Smirnov Test check if the concentrations are normaly distributed
pval_ks_test <- ks.test(x = mean_concentration$Mean, 
        y = "pnorm", 
        mean = mean(mean_concentration$Mean), 
        sd = sd(mean_concentration$Mean))$p.value
pval_ks_test
# Data is normal distributed if the p-value > 0.05. So the concentration variable is NOT normally distributed.
```



# Transform/normalize/scale
Getting a dataframe with log and scaled concentrations
```{r echo=FALSE}
# Logtransform data
metabolite_data_log <- met_df_imputed
metabolite_data_log[,3:ncol(metabolite_data_log)] <- log2(metabolite_data_log[,3:ncol(metabolite_data_log)])

# Scale data 
metabolite_data_log_scale <- metabolite_data_log
metabolite_data_log_scale[,3:ncol(met_df_imputed)] <- scale(metabolite_data_log[,3:ncol(metabolite_data_log)])

# Save data frame 
metabolite_data_log_scale <- as_tibble(metabolite_data_log_scale) %>% 
  mutate(Sample_ID = case_when(nchar(Sample_ID) == 1 ~ as.character(paste0("00", Sample_ID)),
                               nchar(Sample_ID) == 2 ~ as.character(paste0("0", Sample_ID))))  %>% 
  write_csv(paste(path, "data/02_metabolomics_imputed_data_log2_zscore.csv", sep="/")) 
```


## Overview via boxplot
```{r echo=FALSE}
# Pivot data frame 
pivot_metabo_data <- metabolite_data_log_scale %>% 
  pivot_longer(cols = -c(Sample_ID, Label), 
               names_to = "Metabolite", 
               values_to = "Concentrations")

# Boxplot of raw metabolite concentrations
metabolite_plot <- pivot_metabo_data %>% 
  ggplot(mapping = aes(x = Concentrations, y = Metabolite)) + 
  geom_boxplot(alpha = 0.5) + 
  geom_point(position = position_jitter(seed = 0.4, width = 0.02), size = 0.8) +
  coord_flip() + 
  labs(x = 'Imputed, log2, and scaled metabolite concentrations', title = "Distribution of metabolite concentrations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.5))
metabolite_plot
```


# Clinical data: Impute, log transform and scale clinical data
```{r}
clin <- clinical_data_clean %>% 
  mutate(bwcat_binary = case_when(bwcat == "NBW" ~ "0",
                                  bwcat == "LBW" ~ "1", 
                                  TRUE ~ bwcat),
         visit_binary = case_when(visit == "A" ~ "0",
                                  visit == "B" ~ "1", 
                                  visit == "C" ~ "2",
                                  TRUE ~ visit),
         randomization_binary = case_when(randomization == "Control" ~ "0",
                                          randomization == "Training" ~ "1", 
                                          TRUE ~ randomization),
         NAFLD_binary = case_when(NAFLD == "FALSE" ~ "0",
                                  NAFLD == "TRUE" ~ "1"),
         bwcat_NAFLD_binary = case_when(bwcat_NAFLD == "NBW_FALSE" ~ "0",
                                        bwcat_NAFLD == "LBW_FALSE" ~ "1", 
                                        bwcat_NAFLD == "LBW_TRUE" ~ "2",
                                        TRUE ~ bwcat_NAFLD)) %>% 
  select(id, bwcat_binary, visit_binary, NAFLD_binary, bwcat_NAFLD_binary, randomization_binary, everything(), -c(bwcat, visit, NAFLD, bwcat_NAFLD, randomization, visitid)) 
```

```{r}
dim(clin)
```

```{r}
table(is.na(clin))
```


## Impute data with k-nearest neighbor
```{r}
# Convert data frame from numeric to factor
indx <- sapply(clin, is.numeric)
clin[indx] <- lapply(clin[indx], function(clin) as.factor(as.character(clin)))
clin_ <- clin[1:71]

# Set random seed 
if(exists(".Random.seed")) rm(.Random.seed)
library(impute) # Imputing data with KNN

# Impute with KNN function
clin_data_imputed <- impute.knn(as.matrix(clin_)) 
  
clin_data_df_imputed <- as_tibble(clin_data_imputed$data) %>% 
   mutate(across(.cols = 1:AUC_ins, as.numeric))
```



## Transform/normalize/scale
Getting a dataframe with log and scaled concentrations
```{r echo=FALSE}
# Logtransform data
clin_data_log <- clin_data_df_imputed
clin_data_log[,7:ncol(clin_data_log)] <- log2(clin_data_log[,7:ncol(clin_data_log)])

# Scale data 
clin_data_log_scale <- clin_data_log
clin_data_log_scale[,7:ncol(clin_data_log)] <- scale(clin_data_log[,7:ncol(clin_data_log)]) 

clin_data_log_scale_ <- clin_data_log_scale %>%  
  mutate(visit = case_when(visit_binary == 0 ~ "A",
                           visit_binary == 1 ~ "B", 
                           visit_binary == 2 ~ "C"),
         bwcat = case_when(bwcat_binary == 0 ~ "NBW",
                           bwcat_binary == 1 ~ "LBW"), 
         randomization = case_when(randomization_binary == 0 ~ "Control",
                                   randomization_binary == 1 ~ "Training"),
         NAFLD = case_when(NAFLD_binary == 0 ~ "FALSE",
                           NAFLD_binary == 1 ~ "TRUE" ),
         bwcat_NAFLD = case_when(bwcat_NAFLD_binary == 0 ~ "NBW_FALSE",
                                 bwcat_NAFLD_binary == 1 ~ "LBW_FALSE", 
                                 bwcat_NAFLD_binary == 2 ~ "LBW_TRUE"))
  

# Save data frame 
clin_data_log_scale_ <- as_tibble(clin_data_log_scale_) %>% 
  mutate(id = case_when(nchar(id) == 1 ~ as.character(paste0("00", id)),
                        nchar(id) == 2 ~ as.character(paste0("0", id))))  %>% 
  write_csv(paste(path, "data/02_clinical_imputed_data_log2_zscore.csv", sep="/")) 
```





# Combine cleaned and log2/zscore data frames
```{r echo=FALSE}
combined_data <- left_join(metabolite_data_log_scale, clin_data_log_scale_,  by = c("Label" = "visit", "Sample_ID" = "id")) %>% 
  select(Sample_ID, bwcat, visit_binary, NAFLD, randomization, bwcat_NAFLD, bwcat_binary, bwcat_NAFLD_binary, randomization_binary, NAFLD_binary, everything()) %>% 
  write_csv(paste(path, "data/02_combined_data_imputed_log2_zscore.csv", sep="/")) 
combined_data
```

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------