---
title: "Metabolomics Baseline Analysis"
subtitle: "Biological Network Topology Analysis"
author: "Sofie Olund Villumsen"
date: "17/9/2021"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

# Initial settings
```{r echo=FALSE}
# Clear workspace
rm(list = ls()) 
```

Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(correlation) # Compute rho and p-values of the correlation test for several pairs of variables at once
library(igraph)
library(leidenAlg) # Community detection
```

Load data
```{r message=FALSE}
path <- "L:/LovbeskyttetMapper/HCOF Stem Cells/Sofie/Baseline_study/Metabolomics"
setwd(path)

metabolite_data <- read_csv(paste(path, "data/02_metabolomics_imputed_data_log2_zscore.csv", sep="/"))
metabolite_database <- read_csv(paste(path, "data/02_metabolite_database_info.csv", sep="/")) %>% 
  mutate(mmetabolites = sub("^(\\d)", "X\\1", metabolites)) %>% 
  mutate(mmetabolites = str_replace_all(mmetabolites, "[- ,]", "."))

met_clin_data <- read_csv(paste(path, "data/02_combined_data_imputed_log2_zscore.csv", sep="/"))

top_metabolites <- read_csv(paste(path, "data/04_metabolomics_top_metabolites.csv", sep="/"))

fdr_metabolites <- top_metabolites %>%  filter(adj.P.Val <= 0.1) %>% select(metabolites) %>% unique()
pvalue_metabolites <- top_metabolites %>%  filter(adj.P.Val > 0.1) %>% select(metabolites) %>% unique()
```

Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
```{r}
set.seed(422053)
```


```{r echo=FALSE}
# Get a list of metabolites
met_col_names <- colnames(metabolite_data)[3:ncol(metabolite_data)]

metabolite_list <- colnames(metabolite_data)
```

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------


# Data preparation
```{r}
met_df <- metabolite_data %>% 
  filter(Label == "A") 

met_df_2 <- met_df %>% 
  select(-everything(), all_of(met_col_names)) %>% 
  t()

colnames(met_df_2) <- met_df$Sample_ID
rownames(met_df_2) <- met_col_names
met_df_2 <- as_tibble(met_df_2) %>% 
  mutate(metabolites = met_col_names) %>% 
  select(metabolites, everything()) %>% 
  rename(mmetabolites = metabolites)

kegg <- left_join(met_df_2, metabolite_database,"mmetabolites") %>% 
  select(KEGG, HMDB, all_of(colnames(met_df_2)), -c("PubChem", "CheBI")) 

write_csv(kegg, paste(path, "data/05_pathway.csv", sep="/")) 
```


------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

------------------------------------------------------------------------------------------------------------------------------

# Positive + Negative Correlations
## Association Analysis
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1

metabolite_conc <- met_df %>% 
  select(everything(), -c(Label, Sample_ID))

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix <- correlation(metabolite_conc, method = "spearman", p_adjust = "BH") %>%  as_tibble()

corr_matrix_all <- corr_matrix %>%
  rename(Metabolite1 = Parameter1, 
         Metabolite2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  select(Metabolite1, Metabolite2, rho) %>% 
  mutate(Metabolite1 = gsub(",", "_", Metabolite1),
         Metabolite2 = gsub(",", "_", Metabolite2)) %>% 
  as.data.frame()

write_csv(corr_matrix_all, paste(path, "data/05_all_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_all <- graph_from_data_frame(corr_matrix_all, directed = FALSE)
g_all <- set_edge_attr(g_all, "weight", value= corr_matrix_all$rho)
is_weighted(g_all)

# Plot graph
plot(g_all, vertex.label = V(g_all)$name)
```

## Random network 
```{r}
# Number of nodes in positive network
n_nodes <- unique(c(corr_matrix_all$Metabolite1, corr_matrix_all$Metabolite2))

# Graph of random network, unweighted, node and edge number based on a network of the same size
g_random <- erdos.renyi.game(n = length(n_nodes),           # Number of nodes in the graph
                             p.or.m = length(corr_matrix_all$rho), # Number of edges in the graph
                             type = "gnm",
                             directed = FALSE)

# Leidenalg: Community detection algorithm 
leiden_random <- leiden.community(g_random)

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_random, membership(leiden_random))

degree_distribution(g_random)
plot(g_random)
```


## Comparison: pos+neg network
vs random network
```{r}
all_char <- c(gorder(g_all), 
              gsize(g_all),
              edge_density(g_all),
              transitivity(g_all, type = ("undirected")), # global clustering coefficient
              min_cut(g_all)
              )

random_char <- c(gorder(g_random), 
                 gsize(g_random), 
                 edge_density(g_random),
                 transitivity(g_random, type = ("undirected")), # global clustering coefficient
                 min_cut(g_random)
                 )

network_characteristics <- as_tibble(rbind(all_char, random_char)) %>% 
  mutate(network_type = c("pos_neg_network", "random_network")) %>% 
  rename(n_nodes = "V1", n_edges = "V2", density = "V3", cluster_coeff = "V4", min_cut ="V5") %>% 
  select(network_type, everything())
network_characteristics
```

Density = the number of edges a node has, divided by the total possible edges a node could have
....


## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
# Leidenalg: Community detection algorithm 
leiden_community <- leiden.community(g_all)

# Community seperation
g_all_community <- as_tibble(leiden_community$membership)  %>% 
  mutate(Metabolites = leiden_community$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_all, membership(leiden_community))

# Number of communities and how many metabolites each contain
sizes(leiden_community)


##wc <- cluster_walktrap(g_pos)
##plot(wc, g_pos)

##plot(leiden_community$names, g_pos)
```


```{r}
# Create table which contain information on the metabolites
node_table_all <- g_all_community %>% 
  rename(Community_nr = value) %>% 
  select(Metabolites, Community_nr)

node_table_all2 <- left_join(node_table_all, fdr_metabolites, by = c("Metabolites" = "metabolites"), keep = TRUE)
node_table_all3 <- left_join(node_table_all2, pvalue_metabolites, by = c("Metabolites" = "metabolites"), keep = TRUE) %>% 
  rename(fdr_metabolites = metabolites.x, pvalue_metabolites = metabolites.y)

node_table_all3
write_csv(node_table_all3, paste(path, "data/05_all_node_table_imputed.csv", sep="/"))
```

---------------------------------------------------------------------------------------------------------------------------------

---------------------------------------------------------------------------------------------------------------------------------

















# Positive correlations
```{r}
corr_matrix_pos <- corr_matrix_all %>% 
  filter(rho > 0) %>% 
  select(Metabolite1, Metabolite2, rho) %>% 
  as.data.frame()

write_csv(corr_matrix_pos, paste(path, "data/05_pos_edge_table.csv", sep="/"))

# Graph of positive associations, weighted
g_pos <- graph_from_data_frame(corr_matrix_pos, directed = FALSE)
g_pos <- set_edge_attr(g_pos, "weight", value= corr_matrix_pos$rho)
is_weighted(g_pos)

# Plot graph
plot(g_pos, vertex.label = V(g_pos)$name)
```

The above plot shows a biological network with weights (spearmann). A network makes it easier to discover trends/groupings in data and give us a more holistic view of the correlations between the molecules. By applying a community analysis (leiden algorithm) we are able to discover biochemicals with common properties and keep them tightly connected (check among others the modularity score, as a step in the robustness check of the network). 


# Random network
```{r}
# Number of nodes in positive network
n_nodes <- unique(c(corr_matrix_pos$Metabolite1, corr_matrix_pos$Metabolite2))

# Graph of random network, unweighted, node and edge number based on a network of the same size
g_random <- erdos.renyi.game(n = length(n_nodes),           # Number of nodes in the graph
                             p.or.m = length(corr_matrix_pos$rho), # Number of edges in the graph
                             type = "gnm",
                             directed = FALSE)
degree_distribution(g_random)
plot(g_random)
```


# Comparison: positive vs random network
```{r}
pos_char <- c(gorder(g_pos), 
              gsize(g_pos),
              diameter(g_pos, directed = FALSE, unconnected = TRUE, weights = NULL), 
              edge_density(g_pos),
              transitivity(g_pos, type = ("undirected")), # global clustering coefficient
              min_cut(g_pos)
              )

random_char <- c(gorder(g_random), 
                 gsize(g_random), 
                 diameter(g_random, directed = FALSE, unconnected = TRUE, weights = NULL), 
                 edge_density(g_random),
                 transitivity(g_random, type = ("undirected")), # global clustering coefficient
                 min_cut(g_random)
                 )

network_characteristics <- as_tibble(rbind(pos_char, random_char)) %>% 
  mutate(network_type = c("pos_network", "random_network")) %>% 
  rename(n_nodes = "V1", n_edges = "V2", diameter = "V3", density = "V4", cluster_coeff = "V5", min_cut ="V6") %>% 
  select(network_type, everything())
network_characteristics
```

Diameter = a higher diameter indicates that there are some nodes in the perifery (classic for a biological network)
Density = the number of edges a node has, divided by the total possible edges a node could have
....


# Positive graph: Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
# Leidenalg: Community detection algorithm 
leiden_community <- leiden.community(g_pos)

# Community seperation
g_pos_community <- as_tibble(leiden_community$membership)  %>% 
  mutate(Metabolites = leiden_community$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_pos, membership(leiden_community))

# Number of communities and how many metabolites each contain
sizes(leiden_community)


##wc <- cluster_walktrap(g_pos)
##plot(wc, g_pos)

##plot(leiden_community$names, g_pos)
```




```{r}
# Create table which contain information on the metabolites
node_table <- left_join(x = as_tibble(g_pos_community), 
                        y = as_tibble(kegg),
                        kegg$KEGG, by = c("Metabolites" = "metabolites")) %>% 
  rename(Community_nr = value) %>% 
  select(Metabolites, Community_nr, KEGG, HMDB)

node_table_2 <- left_join(node_table, fdr_metabolites, by = c("Metabolites" = "value"), keep = TRUE)
node_table_3 <- left_join(node_table_2, pvalue_metabolites, by = c("Metabolites" = "value"), keep = TRUE) %>% 
  rename(fdr_metabolites = value.x, pvalue_metabolites = value.y) %>% 
  mutate(Metabolites = gsub(",", "_", Metabolites))

node_table_3
write_csv(node_table_3, paste(path, "data/05_node_table.csv", sep="/"))
```


# Random graph
```{r}
# Leidenalg: Community detection algorithm 
leiden_random <- leiden.community(g_random)

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_random, membership(leiden_random))
```



- https://igraph.org/r/doc/ 



-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------

# NBW
```{r}
nbw <- met_clin_data %>% 
  filter(Label == "A", bwcat == "NBW") %>% 
  select(-everything(), all_of(met_col_names))
nbw

nbw <- met_clin_data %>% 
  filter(Label == "A", bwcat == "NBW") %>% 
  select(-c(Sample_ID, bwcat, visit_binary, NAFLD, randomization, bwcat_NAFLD, bwcat_binary, bwcat_NAFLD_binary, randomization_binary, NAFLD_binary, Label))
nbw
```

## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.4

biochem_conc_nbw <- nbw %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_nbw <- correlation(biochem_conc_nbw, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Edge table: Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
corr_matrix_nbw_posneg <- corr_matrix_nbw %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_nbw_posneg, paste(path, "data/06_NBW_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_nbw <- graph_from_data_frame(corr_matrix_nbw_posneg, directed = FALSE)
g_posneg_nbw <- set_edge_attr(g_posneg_nbw, "weight", value= corr_matrix_nbw_posneg$rho)
is_weighted(g_posneg_nbw)

# Plot graph
plot(g_posneg_nbw, vertex.label = V(g_posneg_nbw)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_nbw <- leiden.community(g_posneg_nbw)

# Community seperation
g_posneg_nbw_community <- as_tibble(leiden_community_nbw$membership)  %>% 
  mutate(Biochemicals = leiden_community_nbw$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_nbw, membership(leiden_community_nbw))

# Number of communities and how many metabolites each contain
sizes(leiden_community_nbw)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_nbw_posneg <- g_posneg_nbw_community %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr) %>% 
  mutate(group_met = case_when(Biochemicals %in% metabolite_list ~ "metabolite")) %>% 
  mutate(group_clin = case_when(is.na(group_met) ~ "clinical")) %>% 
  unite(group, group_met:group_clin, remove = TRUE, na.rm = TRUE)

node_table_nbw_posneg2 <- left_join(node_table_nbw_posneg, fdr_metabolites, by = c("Biochemicals" = "metabolites"), keep = TRUE)
node_table_nbw_posneg3 <- left_join(node_table_nbw_posneg2, pvalue_metabolites, by = c("Biochemicals" = "metabolites"), keep = TRUE) 
node_table_nbw_posneg4 <- node_table_nbw_posneg3 %>% 
  rename(fdr_metabolites = metabolites.x, pvalue_metabolites = metabolites.y) %>%
  unite(col = "significant_biochemicals",  fdr_metabolites, pvalue_metabolites, na.rm=TRUE, remove = FALSE) 


node_table_nbw_posneg4
write_csv(node_table_nbw_posneg4, paste(path, "data/06_NBW_posneg_node_table_imputed.csv", sep="/"))
```



-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------

# LBW
```{r}
lbw <- met_clin_data %>% 
  filter(Label == "A", bwcat == "LBW") %>% 
  select(-everything(), all_of(met_col_names))

lbw <- met_clin_data %>% 
  filter(Label == "A", bwcat == "LBW") %>% 
  select(-c(bwcat, Label, NAFLD, bwcat_NAFLD, randomization, visitid, Sample_ID))
lbw
```

## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.4

biochem_conc_lbw <- lbw %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_lbw <- correlation(biochem_conc_lbw, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
corr_matrix_lbw_posneg <- corr_matrix_lbw %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_lbw_posneg, paste(path, "data/06_LBW_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_lbw <- graph_from_data_frame(corr_matrix_lbw_posneg, directed = FALSE)
g_posneg_lbw <- set_edge_attr(g_posneg_lbw, "weight", value= corr_matrix_lbw_posneg$rho)
is_weighted(g_posneg_lbw)

# Plot graph
plot(g_posneg_lbw, vertex.label = V(g_posneg_lbw)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_lbw <- leiden.community(g_posneg_lbw)

# Community seperation
g_posneg_lbw_community <- as_tibble(leiden_community_lbw$membership)  %>% 
  mutate(Biochemicals = leiden_community_lbw$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_lbw, membership(leiden_community_lbw))

# Number of communities and how many metabolites each contain
sizes(leiden_community_lbw)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_lbw_posneg <- g_posneg_lbw_community %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr) %>% 
  mutate(group_met = case_when(Biochemicals %in% metabolite_list ~ "metabolite")) %>% 
  mutate(group_clin = case_when(is.na(group_met) ~ "clinical")) %>% 
  unite(group, group_met:group_clin, remove = TRUE, na.rm = TRUE)

node_table_lbw_posneg2 <- left_join(node_table_lbw_posneg, fdr_metabolites, by = c("Biochemicals" = "metabolites"), keep = TRUE)
node_table_lbw_posneg3 <- left_join(node_table_lbw_posneg2, pvalue_metabolites, by = c("Biochemicals" = "metabolites"), keep = TRUE) 
node_table_lbw_posneg4 <- node_table_lbw_posneg3 %>% 
  rename(fdr_metabolites = metabolites.x, pvalue_metabolites = metabolites.y) %>%
  unite(col = "significant_biochemicals",  fdr_metabolites, pvalue_metabolites, na.rm=TRUE, remove = FALSE) 


node_table_lbw_posneg4
write_csv(node_table_lbw_posneg4, paste(path, "data/06_LBW_posneg_node_table_imputed.csv", sep="/"))
```

-------------------------------------------------------------
 
-------------------------------------------------------------

-------------------------------------------------------------

# LBW-NAFLD
```{r}
lbw_noNAFLD <- met_clin_data %>% 
  filter(Label == "A", bwcat == "LBW", NAFLD == "FALSE") %>% 
  select(-everything(), all_of(met_col_names))

lbw_noNAFLD <- met_clin_data %>% 
  filter(Label == "A", bwcat == "LBW", NAFLD == "FALSE") %>% 
  select(-c(bwcat, Label, NAFLD, bwcat_NAFLD, randomization, visitid, Sample_ID))
lbw_noNAFLD
```

## Data preparation 
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
# LINK: https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/
# LINK github correlation package: https://easystats.github.io/correlation/

# Set variable for the cut-off value of the multiple hypothesis testing (< .001***, < .01**, < .1*)
alpha_thres <- 0.1
rho_thres <- 0.4

biochem_conc_lbw_noNAFLD <- lbw_noNAFLD %>% 
  select(everything())

# Compute correlation (rho) and p-values (BH) for several pairs of variables at once.
corr_matrix_lbw_noNAFLD <- correlation(biochem_conc_lbw_noNAFLD, method = "spearman", p_adjust = "BH") %>%  as_tibble()
```


## Positive + Negative Correlations
The initial network analysis will be an association analysis using Spearmans rank correlations. The network has a big chance of producing false positives, thus the Benjamin-Hochberg (BH) for false discovery rate (FDR).
```{r}
corr_matrix_lbw_noNAFLD_posneg <- corr_matrix_lbw_noNAFLD %>%
  rename(Biochemical1 = Parameter1, 
         Biochemical2 = Parameter2, 
         pval_adj_BH = p) %>% 
  filter(pval_adj_BH < alpha_thres) %>% 
  filter(abs(rho) > rho_thres) %>%  # Threshold of rho is set to be over 0.4 (asolute value of rho)
  select(Biochemical1, Biochemical2, rho) %>% 
  mutate(Biochemical1 = gsub(",", "_", Biochemical1),
         Biochemical2 = gsub(",", "_", Biochemical2)) %>% 
  as.data.frame()

write_csv(corr_matrix_lbw_noNAFLD_posneg, paste(path, "data/06_LBW_noNAFLD_posneg_edge_table_imputed.csv", sep="/"))

# Graph of associations, weighted
g_posneg_lbw_noNAFLD <- graph_from_data_frame(corr_matrix_lbw_noNAFLD_posneg, directed = FALSE)
g_posneg_lbw_noNAFLD <- set_edge_attr(g_posneg_lbw_noNAFLD, "weight", value= corr_matrix_lbw_noNAFLD_posneg$rho)
is_weighted(g_posneg_lbw_noNAFLD)

# Plot graph
plot(g_posneg_lbw_noNAFLD, vertex.label = V(g_posneg_lbw_noNAFLD)$name)
```

## Community detection
Community detection techniques are useful to discover biochemicals with common properties and keep them tightly connected
```{r}
#Seed: Set a seed for reproducibility of data. Seed sampled with -> sample(1e6, 1)
set.seed(758474)

# Leidenalg: Community detection algorithm 
leiden_community_lbw_noNAFLD <- leiden.community(g_posneg_lbw_noNAFLD)

# Community seperation
g_posneg_lbw_noNAFLD_community <- as_tibble(leiden_community_lbw_noNAFLD$membership)  %>% 
  mutate(Biochemicals = leiden_community_lbw_noNAFLD$names) 

# Modularity of network: in a range from -1 to 1, a value close to 1 indicates strong community structure
modularity(g_posneg_lbw_noNAFLD, membership(leiden_community_lbw_noNAFLD))

# Number of communities and how many metabolites each contain
sizes(leiden_community_lbw_noNAFLD)
```


## Node table
```{r}
# Create table which contain information on the metabolites
node_table_lbw_noNAFLD_posneg <- g_posneg_lbw_noNAFLD_community %>% 
  rename(Community_nr = value) %>% 
  select(Biochemicals, Community_nr) %>% 
  mutate(group_met = case_when(Biochemicals %in% metabolite_list ~ "metabolite")) %>% 
  mutate(group_clin = case_when(is.na(group_met) ~ "clinical")) %>% 
  unite(group, group_met:group_clin, remove = TRUE, na.rm = TRUE)

node_table_lbw_noNAFLD_posneg2 <- left_join(node_table_lbw_noNAFLD_posneg, fdr_metabolites, by = c("Biochemicals" = "metabolites"), keep = TRUE)
node_table_lbw_noNAFLD_posneg3 <- left_join(node_table_lbw_noNAFLD_posneg2, pvalue_metabolites, by = c("Biochemicals" = "metabolites"), keep = TRUE) 
node_table_lbw_noNAFLD_posneg4 <- node_table_lbw_noNAFLD_posneg3 %>% 
  rename(fdr_metabolites = metabolites.x, pvalue_metabolites = metabolites.y) %>%
  unite(col = "significant_biochemicals",  fdr_metabolites, pvalue_metabolites, na.rm=TRUE, remove = FALSE) 


node_table_lbw_noNAFLD_posneg4
write_csv(node_table_lbw_noNAFLD_posneg4, paste(path, "data/06_LBW_noNAFLD_posneg_node_table_imputed.csv", sep="/"))
```






